---
title: "Pacbio Phage Analysis"
author: "Eric Z Chen"
date: "September 30, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

Rick: I'm writing to follow up on the possible use of Kmer analysis in analyzing PacBio sequencing data for motifs assocaited with covalent DNA modification.  The PacBio sequencing method involves single DNA molecules traversing along single immobilized polymerase molecules.  The method is sensitive to covalent modification fo DNA, which results in altered kinetics of polymerization. We have been trying to apply this to metagenomic bacteriophage samples.  Phage are known to engage in more than 10 forms of covalent DNA modification. Our metagenomic data is low coverage and noisy.  It might help pull out the signal to search within motifs.  That is, the A of GATC is known to be modified in many organisms.  If we focused on the A in GATC, we would see a strong signal in some contigs.  Modification is often sequence specific, but we do not always know the sequences involved in metagenomic samples. Thus it would be great to test all 2mers, 3mers…etc against all contigs, and tabulate each base within each Nmer to check for altered kinetics.  We would also test data on control DNA where we are sure there are no mutations as a control.  We would tabulate those motifs with altered kinetics, and show that these were rare in control data.  This would provide a set of new specificities for DNA modifying enzymes encoded in phage genomes, data which could find many uses.
```{r setup, results='hide',warning=FALSE,message=FALSE}
library(knitr)
library(mixtools)
library(stringr)
library(ggplot2)
#devtools::install_github("chvlyl/metakit")
library(metakit)
library(pROC)
library(readr)
library(randomForest)
library(tidyr)
library(dplyr)
```

```{r note, eval=FALSE}
1. Include the second bacterial data
2. Remove the repeat sequences (lower letter) 
3. Remove short contig
4. Take a log transformation
5. Fit Laplacian mixture model

Remove sequences at the begnning and the end
Remove short contigs
Remove contigs with high GC>95%: unitig_336
```


### Raw data
```{r load_data,cache=TRUE,eval=TRUE,cache.lazy=FALSE}
#################
### Error: "large vector not supported"
### update the knitr and rmarkdown packages
#################

###################################
### load control T4 phage sequences
t4.phage.seq <- read_tsv(file='../0_Data/Processed_Data/1_T4_phage_all.fasta.table', col_names = c('ID','Seq'))
apply(t4.phage.seq[,2],1,nchar)
## [1] 168922 166816 165660 168922
## The four phage genomes are different
## Use their correpsonding genome sequence in the analysis

### load control T4 phage IPD
T4GlcHMC.ipd <- read_tsv(file='../0_Data/Processed_Data/1_T4_Phage_GlcHMC.txt', col_names = TRUE)
T4HMC.ipd <- read_tsv(file='../0_Data/Processed_Data/2_T4_Phage147_HMC.txt', col_names = TRUE)
T4C.ipd <- read_tsv(file='../0_Data/Processed_Data/3_T4_PhageGT7_C.txt', col_names = TRUE)
T4WGA.ipd <- read_tsv(file='../0_Data/Processed_Data/4_T4_Phage_WGA.txt', col_names = TRUE)

###################
### load phage IPD
###  [1] "refName"         "tpl"             "strand"          "base"       ### [5] "score"           "tMean"           "tErr"            "modelPrediction"
### [9] "ipdRatio"        "coverage"        "frac"            "fracLow"     ### [13] "fracUp"  

#########
## 0 is Fwd strand


## phage 1
phage1.seq <- read_tsv(file='../0_Data/Processed_Data/5_Phage1.table', col_names = c('ID','Seq'))
apply(phage1.seq[,2],1,nchar) %>% sort
  
## phage 1
phage1.ipd <- read_csv(file='../0_Data/Processed_Data/5_Phage1.csv', col_names = TRUE)
phage1.ipd <- phage1.ipd %>%
dplyr::rename(ID=refName,Pos=tpl,Strand=strand,Base=base,Score=score,
              IPDRatio=ipdRatio,Cov=coverage) %>%
dplyr::mutate(Strand=ifelse(Strand==0,'Fwd','Rev')) %>%
dplyr::mutate(Type='Phage1') %>%
dplyr::select(Type, ID, Pos, Strand, Base, Cov, IPDRatio, Score) 


## phage 2
phage2.ipd <- read_csv(file='../0_Data/Processed_Data/6_Phage2.csv', col_names = TRUE)
phage2.ipd <- phage2.ipd %>%
dplyr::rename(ID=refName,Pos=tpl,Strand=strand,Base=base,Score=score,
              IPDRatio=ipdRatio,Cov=coverage) %>%
dplyr::mutate(Strand=ifelse(Strand==0,'Fwd','Rev')) %>%
dplyr::mutate(Type='Phage2') %>%
dplyr::select(Type, ID, Pos, Strand, Base, Cov, IPDRatio, Score)


### bact 1
bact1.ipd <- read_csv(file='../0_Data/Processed_Data/7_Bacteria1.csv', col_names = TRUE)
bact1.ipd <- bact1.ipd %>%
dplyr::rename(ID=refName,Pos=tpl,Strand=strand,Base=base,Score=score,
              IPDRatio=ipdRatio,Cov=coverage) %>%
dplyr::mutate(Strand=ifelse(Strand==0,'Fwd','Rev')) %>%
dplyr::mutate(Type='Bact1') %>%
dplyr::select(Type, ID, Pos, Strand, Base, Cov, IPDRatio, Score)
 
### bact 2
bact2.ipd <- read_csv(file='../0_Data/Processed_Data/8_Bacteria2.csv', col_names = TRUE)
bact2.ipd <- bact2.ipd %>%
dplyr::rename(ID=refName,Pos=tpl,Strand=strand,Base=base,Score=score,
              IPDRatio=ipdRatio,Cov=coverage) %>%
dplyr::mutate(Strand=ifelse(Strand==0,'Fwd','Rev')) %>%
dplyr::mutate(Type='Bact2') %>%
dplyr::select(Type, ID, Pos, Strand, Base, Cov, IPDRatio, Score)

#################
### combine T4 IPD
ipd.raw <- rbind(
data.frame(Type='T4', ID='T4GlcHMC',T4GlcHMC.ipd,stringsAsFactors =FALSE),
data.frame(Type='T4',ID='T4HMC',T4HMC.ipd,stringsAsFactors =FALSE),
data.frame(Type='T4',ID='T4C',T4C.ipd,stringsAsFactors =FALSE),
## negative control
data.frame(Type='T4',ID='T4WGA',T4WGA.ipd,stringsAsFactors =FALSE)
)
 
### transformation
ipd <- ipd.raw %>%
tidyr::gather(Var,Value,-Type,-ID,-Pos) %>%
tidyr::separate(Var, into = c("Varable", "Strand"), sep = "_") %>%
tidyr::spread(Varable,Value,convert=TRUE)  %>%
dplyr::mutate(IPDRatio=abs(as.numeric(IPDRatio)))

### check column names
identical(colnames(ipd),colnames(phage1.ipd))
identical(colnames(ipd),colnames(phage2.ipd))
identical(colnames(ipd),colnames(bact1.ipd))
identical(colnames(ipd),colnames(bact2.ipd))
##Type,ID,Pos,Strand,Base,Cov,IPDRatio,Score

### rbind does not work when generating html file
ipd <- rbind(ipd, phage1.ipd, phage2.ipd, bact1.ipd,bact2.ipd)
write_tsv(ipd,path='../0_Data/Merged_IPD/ipd_merged.txt')
```



```{r check_outliers}

## C in GlcHMC are all modified
## WGA no modification

## T4C
ipd %>% dplyr::filter(Type=='T4C') %>%
#dplyr::filter(Strand=='Fwd') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)

## T4WGA
ipd %>% dplyr::filter(Type=='T4WGA') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)

## T4GlcHMC
ipd %>% dplyr::filter(Type=='T4GlcHMC') %>%
#dplyr::filter(Strand=='Fwd') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)

## T4HMC
ipd %>% dplyr::filter(Type=='T4HMC') %>%
#dplyr::filter(Strand=='Fwd') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)


## Phage1
ipd %>% dplyr::filter(Type=='Phage1') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)


ipd %>% dplyr::filter(Type=='Phage1') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base) + xlim(0,5)


ipd %>% dplyr::filter(Type=='Phage1') %>%
ggplot( aes(log10(IPDRatio), fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)

## Phage2
ipd %>% dplyr::filter(Type=='Phage2') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)

ipd %>% dplyr::filter(Type=='Phage2') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base) + xlim(0,5)

ipd %>% dplyr::filter(Type=='Phage2') %>%
ggplot( aes(log10(IPDRatio), fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)


## Bact1
ipd %>% dplyr::filter(Type=='Bact1') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)

ipd %>% dplyr::filter(Type=='Bact1') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base) + xlim(0,5)

ipd %>% dplyr::filter(Type=='Bact1') %>%
ggplot( aes(log10(IPDRatio), fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)


## all
ipd %>% dplyr::filter(ID=='T4') %>%
#dplyr::filter(Strand=='Fwd') %>%
ggplot( aes(IPDRatio, fill = Type)) +geom_density(alpha = 0.2) +
xlim(0,5) + facet_grid(Strand~ Base)

## log transformation
ipd %>% dplyr::filter(ID=='T4') %>%
#dplyr::filter(Strand=='Fwd') %>%
ggplot( aes(log10(IPDRatio), fill = Type)) +geom_density(alpha = 0.2) +
facet_grid(Strand~ Base)



############
## Outliers
## 1. At the beginning and end of the sequence
#############


## T4WGA
ipd %>% dplyr::filter(Type=='T4WGA') %>%
ggplot( aes(IPDRatio, fill = Base)) + geom_bar(binwidth = 0.1, position="dodge")

ipd %>% dplyr::filter(Type=='T4WGA') %>%
top_n(20, IPDRatio)

outlier.pos <- ipd %>% dplyr::filter(Type=='T4WGA') %>%
top_n(20, IPDRatio) %>%
dplyr::select(Pos) 


t4.phage.seq[4,2] %>%
str_sub(start=outlier.pos$Pos-3,end=outlier.pos$Pos+3)
## outliers at the beginning and the end of the sequence

######
## T4C
ipd %>% dplyr::filter(Type=='T4C') %>%
ggplot( aes(IPDRatio, fill = Base)) + geom_bar(binwidth = 0.1, position="dodge")

ipd %>% dplyr::filter(Type=='T4C') %>%
top_n(20, IPDRatio) 

ipd %>% dplyr::filter(Type=='T4C' & Base=='C') %>%
top_n(20, IPDRatio) 


outlier.pos <- ipd %>% dplyr::filter(Type=='T4C') %>%
top_n(20, IPDRatio) %>%
dplyr::select(Pos) 


t4.phage.seq[3,2] %>%
str_sub(start=outlier.pos$Pos,end=outlier.pos$Pos+6)


######
## T4HMC
ipd %>% dplyr::filter(Type=='T4HMC') %>%
ggplot( aes(IPDRatio, fill = Base)) +geom_density(alpha = 0.2)

ipd %>% dplyr::filter(Type=='T4HMC') %>%
top_n(20, IPDRatio) 

outlier.pos <- ipd %>% dplyr::filter(Type=='T4HMC') %>%
top_n(20, IPDRatio) %>%
dplyr::select(Pos) 


t4.phage.seq[2,2] %>%
str_sub(start=outlier.pos$Pos-2,end=outlier.pos$Pos+6)


######
## T4GlcHMC

ipd %>% dplyr::filter(Type=='T4GlcHMC') %>%
ggplot( aes(IPDRatio, fill = Base)) +geom_density(alpha = 0.2)

ipd %>% dplyr::filter(Type=='T4GlcHMC') %>%
top_n(20, IPDRatio) 


outlier.pos <- ipd %>% dplyr::filter(Type=='T4GlcHMC') %>%
top_n(20, IPDRatio) %>%
dplyr::select(Pos) 


t4.phage.seq[1,2] %>%
str_sub(start=outlier.pos$Pos,end=outlier.pos$Pos+6)


######
## phage1

ipd %>% dplyr::filter(Type=='Phage1') %>%
ggplot( aes(IPDRatio, fill = Base)) +geom_density(alpha = 0.2)

ipd %>% dplyr::filter(Type=='Phage1') %>%
ggplot( aes(IPDRatio, fill = Base)) +geom_density(alpha = 0.2)+
xlim(0,5)

ipd %>% dplyr::filter(Type=='Phage1') %>%
dplyr::filter(IPDRatio>2) %>%
ggplot( aes(IPDRatio, fill = Base)) +geom_density(alpha = 0.2)

ipd %>% dplyr::filter(Type=='Phage1') %>%
ggplot( aes(log10(IPDRatio), fill = Base)) +geom_density(alpha = 0.2)


ipd %>% dplyr::filter(Type=='Phage1') %>%
top_n(20, IPDRatio) 


ipd %>% dplyr::filter(Type=='Phage1') %>%
top_n(20, IPDRatio) 

phage1.seq %>% dplyr::mutate(len=nchar(Seq)) %>% dplyr::select(-Seq) %>%
dplyr::filter(ID=='unitig_984|q3')

phage1.seq[phage1.seq$ID=='unitig_488|q3','Seq'] %>% nchar

phage1.seq[phage1.seq$ID=='unitig_488|q3','Seq'] %>%
str_sub(start=602,end=602)

phage1.seq[phage1.seq$ID=='unitig_984|q3','Seq'] %>%
str_sub(start=307-10,end=307+10)


```



```{r  delete_temporary_data}
rm(T4GlcHMC.ipd)
rm(T4HMC.ipd)
rm(T4C.ipd)
rm(T4WGA.ipd)
rm(phage1.ipd)
rm(phage2.ipd)
rm(bact1.ipd)
rm(bact2.ipd)
rm(ipd.raw)
```



```{r plot_data,eval=FALSE}
plot_ipd = function(pdata){
### IPD by strand
plot(ggplot(pdata, aes(IPDRatio, fill = Strand)) + geom_density(alpha = 0.2))
plot(ggplot(pdata, aes(IPDRatio, fill = Strand)) + geom_bar(binwidth = 0.1, position="dodge"))
plot(ggplot(pdata, aes(y=IPDRatio, x = Strand)) + geom_boxplot())
### IPD by base
plot(ggplot(pdata, aes(IPDRatio, fill = Base)) + geom_density(alpha = 0.2))
plot(ggplot(pdata, aes(IPDRatio, fill = Base)) + geom_bar(binwidth = 0.1, position="dodge"))
plot(ggplot(pdata, aes(y=IPDRatio, x = Base)) + geom_boxplot())
### Cov by strand
plot(ggplot(pdata, aes(Cov, fill = Strand)) + geom_density(alpha = 0.2))
plot(ggplot(pdata, aes(Cov, fill = Strand)) + geom_bar(binwidth = 0.1, position="dodge"))
plot(ggplot(pdata, aes(y=Cov, x = Strand)) + geom_boxplot())
### Cov by base
plot(ggplot(pdata, aes(Cov, fill = Base)) + geom_density(alpha = 0.2))
plot(ggplot(pdata, aes(Cov, fill = Base)) + geom_bar(binwidth = 0.1, position="dodge"))
plot(ggplot(pdata, aes(y=Cov, x = Base)) + geom_boxplot())
### Cov IPD
plot(ggplot(pdata, aes(Cov, IPDRatio,color=Strand)) + geom_point())
plot(ggplot(pdata, aes(Cov, IPDRatio,color=Base)) + geom_point())
### Score IPD
plot(ggplot(pdata, aes(Score, IPDRatio,color=Strand)) + geom_point())
plot(ggplot(pdata, aes(Score, IPDRatio,color=Base)) + geom_point())
}


######################################
### WGA
ipd %>% dplyr::filter(Type=='T4WGA') %>%
plot_ipd()

######################################
### HMC
ipd %>% dplyr::filter(ID=='T4HMC') %>%
plot_ipd

######################################
### GlcHMC
ipd %>% dplyr::filter(ID=='T4GlcHMC') %>%
plot_ipd

######################################
### C
ipd %>% dplyr::filter(ID=='T4C') %>%
plot_ipd

####################################
### compare four T4 phages
ipd  %>% dplyr::filter(Type=='T4') %>%
(function(pdata){
### IPD by type and base
plot(ggplot(pdata, aes(IPDRatio, fill = ID)) + geom_density(alpha = 0.2) + facet_grid(. ~ Base))
plot(ggplot(pdata, aes(IPDRatio, fill = ID)) + geom_bar(binwidth = 0.1, position="dodge")+ facet_grid(. ~ Base))
plot(ggplot(pdata, aes(y=IPDRatio, x = ID)) + geom_boxplot()+ facet_grid(. ~ Base))
})
 
### C in T4C vs. T4WGA
ipd  %>% dplyr::filter(Base=='C' & (ID=='T4C'|ID=='T4WGA')) %>%
(function(pdata){
### IPD by type and base
plot(ggplot(pdata, aes(IPDRatio, fill = ID)) + geom_density(alpha = 0.2) + facet_grid(. ~ Base))
plot(ggplot(pdata, aes(IPDRatio, fill = ID)) + geom_bar(binwidth = 0.1, position="dodge")+ facet_grid(. ~ Base))
plot(ggplot(pdata, aes(y=IPDRatio, x = ID)) + geom_boxplot()+ facet_grid(. ~ Base))
### Cov IPD
plot(ggplot(pdata, aes(Cov, IPDRatio)) + geom_point()+ facet_grid(. ~ Base))
})

### large IPDRatio in T4WGA
ipd  %>% dplyr::filter(ID=='T4WGA' & IPDRatio>2) 


### large IPDRatio in T4WGA
### qqplot 
ipd  %>% dplyr::filter(ID=='T4WGA') %>%
(function(pdata){
plot(ggplot(pdata,aes(sample=IPDRatio))+stat_qq(distribution = qnorm)+ facet_grid(. ~ Base)+ggtitle('Nomral'))
plot(ggplot(pdata,aes(sample=IPDRatio))+stat_qq(distribution = qlnorm)+ facet_grid(. ~ Base)+ggtitle('LogNomral'))
})

### A in T4C
ipd  %>% dplyr::filter(Base=='A' & ID=='T4C') %>%
(function(pdata){
### IPD by type and base
plot(ggplot(pdata, aes(IPDRatio, fill = ID)) + geom_density(alpha = 0.2) + facet_grid(. ~ Base))
plot(ggplot(pdata, aes(log10(IPDRatio), fill = ID)) + geom_density(alpha = 0.2) + facet_grid(. ~ Base))
plot(ggplot(pdata, aes(IPDRatio, fill = ID)) + geom_bar(binwidth = 0.1, position="dodge")+ facet_grid(. ~ Base))
plot(ggplot(pdata, aes(y=IPDRatio, x = ID)) + geom_boxplot()+ facet_grid(. ~ Base))
### Cov IPD
plot(ggplot(pdata, aes(Cov, IPDRatio)) + geom_point()+ facet_grid(. ~ Base))
})



#########################################################
### all by type and base
ipd %>% dplyr::select(Type,Base,Strand,IPDRatio) %>%
(function(pdata){
### raw in (0,10)
plot(ggplot(pdata, aes(IPDRatio, fill = Strand)) + geom_density(alpha = 0.2)+ facet_grid(Type ~ Base)+xlim(0,10))
### log10
plot(ggplot(pdata, aes(log10(IPDRatio), fill = Strand)) + geom_density(alpha = 0.2)+ facet_grid(Type ~ Base))
})

### large values type and base
ipd %>% dplyr::filter(IPDRatio > 3) %>%
dplyr::select(Type,Base,Strand,IPDRatio) %>%
(function(pdata){
### raw in (0,10)
plot(ggplot(pdata, aes(IPDRatio, fill = Strand)) + geom_density(alpha = 0.2)+ facet_grid(Type ~ Base)+xlim(0,10))
### log10
plot(ggplot(pdata, aes(log10(IPDRatio), fill = Strand)) + geom_density(alpha = 0.2)+ facet_grid(Type ~ Base))
})
```


```{r phage_t4_gatc_in_c_vs_wga,eval=FALSE}
extract_ipd = function(ipd, id,  match.pos, match.base=NULL, str='Fwd'){
  ## "ID"       "Pos"      "Strand"   "Base"     "Cov"      "IPDRatio" "Score"
  ipd.select <- 
  ipd %>% dplyr::filter(ID==id) %>%
  dplyr::filter(Pos %in% match.pos) %>%
  dplyr::filter(Strand==str) %>% 
  dplyr::select(IPDRatio, Cov, Score)
  return(ipd.select)
}

### Positive and negative controls have the same sequence
match.pos <- str_locate_all(pattern ='GATC',ctrl.seq$Seq[1])[[1]][,'start']

### positive controls
g.ipd = extract_ipd(ipd=ipd, id='T4C', match.pos=match.pos, str='Fwd')
a.ipd = extract_ipd(ipd=ipd, id='T4C', match.pos=match.pos+1, str='Fwd')
t.ipd = extract_ipd(ipd=ipd, id='T4C', match.pos=match.pos+2, str='Fwd')
c.ipd = extract_ipd(ipd=ipd, id='T4C', match.pos=match.pos+3, str='Fwd')
t4c.gatc <- data.frame(g=g.ipd,a=a.ipd,t=t.ipd,c=c.ipd,stringsAsFactors =FALSE)

### negative controls
g.ipd = extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos, str='Fwd')
a.ipd = extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+1, str='Fwd')
t.ipd = extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+2, str='Fwd')
c.ipd = extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+3, str='Fwd')
t4wga.gatc <- data.frame(g=g.ipd,a=a.ipd,t=t.ipd,c=c.ipd,stringsAsFactors =FALSE)

### combine them together
gatc.features <- rbind(
cbind(type='pos.ctrl',t4c.gatc),
cbind(type='neg.ctrl',t4wga.gatc)
)


### random forest
cutoff <- 0.6
rf.gatc <- random_forest(gatc.features[,colnames(gatc.features)!='type'], 
               gatc.features[,'type',drop=FALSE],
               plot.MDS=TRUE,plot.importance=TRUE,print.result=TRUE,
               guess.num=0,boxplot.top=12,cutoff=c(cutoff,1-cutoff),
               title="A in GATC",boxplot.y="",ntree=1000,
               seed=10)

```


```{r phage_t4_gatc_in_glchmc_vs_wga,eval=FALSE}
extract_ipd = function(ipd, id,  match.pos, match.base=NULL, str='Fwd'){
  ## "ID"       "Pos"      "Strand"   "Base"     "Cov"      "IPDRatio" "Score"
  ipd.select <- 
  ipd %>% dplyr::filter(ID==id) %>%
  dplyr::filter(Pos %in% match.pos) %>%
  dplyr::filter(Strand==str) %>% 
  dplyr::select(IPDRatio, Cov, Score)
  return(ipd.select)
}

### Positive and negative controls have the same sequence
match.pos <- str_locate_all(pattern ='GATC',ctrl.seq$Seq[1])[[1]][,'start']

### positive controls
g.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos+3, str='Rev')
        )
a.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos+1, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos+2, str='Rev')
        )
t.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos+2, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos+1, str='Rev')
        )
c.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos+3, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4GlcHMC', match.pos=match.pos, str='Rev')
        )
t4c.gatc <- data.frame(g=g.ipd,a=a.ipd,t=t.ipd,c=c.ipd,stringsAsFactors =FALSE)

### negative controls
g.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+3, str='Rev')
        )
a.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+1, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+2, str='Rev')
        )
t.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+2, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+1, str='Rev')
        )
c.ipd = rbind(
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos+3, str='Fwd'),
        extract_ipd(ipd=ipd, id='T4WGA', match.pos=match.pos, str='Rev')
        )
t4wga.gatc <- data.frame(g=g.ipd,a=a.ipd,t=t.ipd,c=c.ipd,stringsAsFactors =FALSE)

### combine them together
gatc.features <- rbind(
cbind(type='pos.ctrl',t4c.gatc),
cbind(type='neg.ctrl',t4wga.gatc)
)


### random forest
cutoff <- 0.6
rf.a.in.gatc.glchmc <- random_forest(gatc.features[,colnames(gatc.features)!='type'], 
               gatc.features[,'type',drop=FALSE],
               plot.MDS=TRUE,plot.importance=TRUE,print.result=TRUE,
               guess.num=0,boxplot.top=12,cutoff=c(cutoff,1-cutoff),
               title="A-in-GATC T4GlcHMC vs. T4WGA",boxplot.y="",ntree=1000,
               seed=10)

####
roc.data <- data.frame(Label=rf.a.in.gatc.glchmc$y,rf.a.in.gatc.glchmc$votes[,1,drop=F])  %>%
dplyr::rename(RandomForest=pos.ctrl) %>%
dplyr::mutate(Label=ifelse(Label=='pos.ctrl',1,0))

roc<-plot.roc(roc.data[,"Label",drop=TRUE], 
              roc.data[,"RandomForest",drop=TRUE], 
         main="A-in-GATC T4C vs. T4WGA", 
         percent=TRUE, col="#e41a1c",type="n",
         print.auc=TRUE,print.auc.pattern="RandomForest AUC:%.1f%%",
         #print.thres=TRUE,
         print.auc.col="#e41a1c",print.auc.y=40)
lines(smooth(roc, method = "binormal"), col="#e41a1c")
legend("bottomright", legend = c("NormalMix"), 
       col = c("#e41a1c"),lwd = 2)

```



```{r fit_mix_model_normal,eval=FALSE}
ipd.filter <- read_tsv(file='../0_Data/Merged_IPD/ipd_merged_filter.txt', col_names = TRUE)


fit_normal_mixture=function(Phage1.ipd){
  Phage1.ipd <- data.frame(Index=1:nrow(Phage1.ipd),Phage1.ipd)
  ##
  a.ipd <- Phage1.ipd  %>% dplyr::filter(Base=='A')  %>% dplyr::select(Index,IPDRatio)
  cutoff <- a.ipd %>% top_n(10,IPDRatio) %>% select(IPDRatio) %>% min
  a.ipd <- a.ipd %>% dplyr::filter(IPDRatio < cutoff)
  a.mfit <- normalmixEM(a.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
  a.ipd  <- a.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(a.ipd, .)
  summary(a.mfit)
  plot(a.mfit,which=2)
  ##
  t.ipd <- Phage1.ipd %>% dplyr::filter(Base=='T')  %>% dplyr::select(Index,IPDRatio)
  cutoff <- t.ipd %>% top_n(10,IPDRatio) %>% select(IPDRatio) %>% min
  t.ipd <- t.ipd %>% dplyr::filter(IPDRatio < cutoff)
  t.mfit <- normalmixEM(t.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
  t.ipd  <- t.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(t.ipd, .)
  summary(t.mfit)
  plot(t.mfit,which=2)
  ##
  c.ipd <- Phage1.ipd  %>% dplyr::filter(Base=='C')  %>% dplyr::select(Index,IPDRatio)
  cutoff <- c.ipd %>% top_n(10,IPDRatio) %>% select(IPDRatio) %>% min
  c.ipd <- c.ipd %>% dplyr::filter(IPDRatio < cutoff)
  c.mfit <- normalmixEM(c.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
  c.ipd  <- c.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(c.ipd, .)
  summary(c.mfit)
  plot(c.mfit,which=2)
  ##
  g.ipd <- Phage1.ipd  %>% dplyr::filter(Base=='G')  %>% dplyr::select(Index,IPDRatio)
  cutoff <- g.ipd %>% top_n(10,IPDRatio) %>% select(IPDRatio) %>% min
  g.ipd <- g.ipd %>% dplyr::filter(IPDRatio < cutoff)
  g.mfit <- normalmixEM(g.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
  g.ipd  <- g.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(g.ipd, .)
  summary(g.mfit)
  plot(g.mfit,which=2)
  ###
  atcg.ipd <- rbind(a.ipd,t.ipd,c.ipd,g.ipd)
  identical(nrow(Phage1.ipd),nrow(atcg.ipd))
  ###
  Phage1.ipd <- Phage1.ipd %>% 
    left_join(atcg.ipd%>%dplyr::select(-IPDRatio),by=c('Index'='Index')) %>% 
    dplyr::rename(unmodified.norm=unmodified,modified.norm=modified)
  
  
  rm(a.ipd)
  rm(t.ipd)
  rm(g.ipd)
  rm(c.ipd)
  rm(atcg.ipd)
  rm(a.mfit)
  rm(t.mfit)
  rm(g.mfit)
  rm(c.mfit)
  
  return(Phage1.ipd)
}


### phage1
T4.ipd <- ipd.filter %>% dplyr::filter(Type=='T4') 
write_tsv(fit_normal_mixture(T4.ipd),
          path='../0_Data/Normal_Mixture_Model/Normal_Mixeture_Model_Fitting_Result_T4.txt')

Phage1.ipd <- ipd.filter %>% dplyr::filter(Type=='Phage1') 
write_tsv(fit_normal_mixture(Phage1.ipd),
          path='../0_Data/Normal_Mixture_Model/Normal_Mixeture_Model_Fitting_Result_Phage1.txt')
Phage2.ipd <- ipd.filter %>% dplyr::filter(Type=='Phage2') 
write_tsv(fit_normal_mixture(Phage2.ipd),
          path='../0_Data/Normal_Mixture_Model/Normal_Mixeture_Model_Fitting_Result_Phage2.txt')
Bact1.ipd <- ipd.filter %>% dplyr::filter(Type=='Bact1') 
write_tsv(fit_normal_mixture(Bact1.ipd),
          path='../0_Data/Normal_Mixture_Model/Normal_Mixeture_Model_Fitting_Result_Bact1.txt')
Bact2.ipd <- ipd.filter %>% dplyr::filter(Type=='Bact2') 
write_tsv(fit_normal_mixture(Bact2.ipd),
          path='../0_Data/Normal_Mixture_Model/Normal_Mixeture_Model_Fitting_Result_Bact2.txt')
```


```{r fit_mix_model_log_normal,eval=FALSE}
### zeros in the IPDRatio
### set them to unmodified

ipd.mix <- data.frame(Index=1:nrow(ipd),ipd)

### phage1
a.ipd <- ipd.mix  %>% dplyr::filter(Type=='Phage1' & Base=='A') %>%
  #dplyr::filter(IPDRatio<5) %>%
  #dplyr::filter(IPDRatio>0) %>%
  #dplyr::mutate(IPDRatio=log(IPDRatio)) %>%  
dplyr::select(Index,IPDRatio)
a.ipd %>% ggplot(aes(IPDRatio)) +geom_density(alpha = 0.2)
a.mfit <- normalmixEM(a.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
a.ipd  <- a.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(a.ipd, .)
summary(a.mfit)


plot(a.mfit,which=2)


t.ipd <- ipd.mix  %>% dplyr::filter(Base=='T') %>% dplyr::mutate(IPDRatio=log10(IPDRatio)) %>% dplyr::select(Index,IPDRatio)
t.mfit <- normalmixEM(t.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
t.ipd  <- t.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(t.ipd, .)
summary(t.mfit)

c.ipd <- ipd.mix  %>% dplyr::filter(Type=='Phage1' &Base=='C') %>%
  dplyr::filter(IPDRatio<5) %>%
  #dplyr::filter(IPDRatio>0) %>%
  #dplyr::mutate(IPDRatio=log10(IPDRatio)) %>% 
  dplyr::select(Index,IPDRatio)
c.mfit <- normalmixEM(c.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
c.ipd  <- c.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(c.ipd, .)
summary(c.mfit)

plot(c.mfit,which=2)

g.ipd <- ipd.mix  %>% dplyr::filter(Base=='G') %>% dplyr::mutate(IPDRatio=log10(IPDRatio)) %>% dplyr::select(Index,IPDRatio)
g.mfit <- normalmixEM(g.ipd$IPDRatio,lambda=0.5,mu=c(1,4),sigma=c(1,2))
g.ipd  <- g.mfit$posterior %>%  as.data.frame %>% dplyr::rename(unmodified=comp.1,modified=comp.2) %>% data.frame(g.ipd, .)
summary(g.mfit)

###
atcg.ipd <- rbind(a.ipd,t.ipd,c.ipd,g.ipd)
identical(nrow(ipd.mix),nrow(atcg.ipd))
ipd.mix <- ipd.mix %>% 
           left_join(atcg.ipd%>%dplyr::select(-IPDRatio),by=c('Index'='Index')) %>% 
           dplyr::select(-Index) %>%
           dplyr::rename(unmodified.lognorm=unmodified,modified.lognorm=modified)

rm(a.ipd)
rm(t.ipd)
rm(g.ipd)
rm(c.ipd)
rm(atcg.ipd)
rm(a.mfit)
rm(t.mfit)
rm(g.mfit)
rm(c.mfit)
```


```{r }
### Positive and negative controls have the same sequence
match.pos <- str_locate_all(pattern ='GATC',ctrl.seq$Seq[1])[[1]][,'start']
a.pos <- sort(c(match.pos+1,match.pos+2))
  
### A in GATC fwd = T in GATC rev
a.plot <- ipd.mix %>% dplyr::filter(ID=='T4GlcHMC' | ID=='T4WGA') %>%
dplyr::filter(Base=='A') %>%
dplyr::mutate(Label=ifelse(ID=='T4GlcHMC',1,0)) %>%
dplyr::filter(Pos %in% a.pos) %>%
dplyr::select(modified.norm,Label)

roc<-plot.roc(a.plot[,"Label",drop=TRUE], 
              a.plot[,"modified.norm",drop=TRUE], 
         main="A-in-GATC T4GlcHMC vs. T4WGA", 
         percent=TRUE, col="#e41a1c",type="n",
         print.auc=TRUE,print.auc.pattern="NormalMix AUC:%.1f%%",
         #print.thres=TRUE,
         print.auc.col="#e41a1c",print.auc.y=40)
lines(smooth(roc, method = "binormal"), col="#e41a1c")
legend("bottomright", legend = c("NormalMix"), 
       col = c("#e41a1c"),lwd = 2)

rm(a.plot)
rm(match.pos)
rm(a.pos)
```


```{r }
### Positive and negative controls have the same sequence
match.pos <- str_locate_all(pattern ='GATC',ctrl.seq$Seq[1])[[1]][,'start']
a.pos <- sort(c(match.pos+1,match.pos+2))
  
### A in GATC fwd = T in GATC rev
a.plot <- ipd.mix %>% dplyr::filter(ID=='T4C' | ID=='T4WGA') %>%
dplyr::filter(Base=='A') %>%
dplyr::mutate(Label=ifelse(ID=='T4C',1,0)) %>%
dplyr::filter(Pos %in% a.pos) %>%
dplyr::select(modified.norm,Label)

  
roc<-plot.roc(a.plot[,"Label",drop=TRUE], 
              a.plot[,"modified.norm",drop=TRUE], 
         main="A-in-GATC T4C vs. T4WGA", 
         percent=TRUE, col="#e41a1c",type="n",
         print.auc=TRUE,print.auc.pattern="NormalMix AUC:%.1f%%",
         #print.thres=TRUE,
         print.auc.col="#e41a1c",print.auc.y=40)
lines(smooth(roc, method = "binormal"), col="#e41a1c")
legend("bottomright", legend = c("NormalMix"), 
       col = c("#e41a1c"),lwd = 2)

rm(a.plot)
rm(match.pos)
rm(a.pos)
```